---
title: "Store Profitability ML Project Proposal"
author: "Shubha S Singh"
format:
  html:
    code-tools: false
    code-fold: false
    toc: true
    toc-depth: 2
execute:
    echo: false
---

# Executive Summary

This project develops machine learning models to improve operations in **Albertsons' meat and seafood Butcher Blocks**. With perishable goods presenting unique challenges around inventory waste and profitability, understanding the relationships between sales performance, shrink rates, and operational efficiency is key to running successful departments.

## Project Objectives

The main goal of this analysis is to **develop a recommendation sheet for butcher block lengths**, determining whether they should be decreased based on each store’s performance.

To support this the analysis will also:

-   Segment stores into clear performance tiers.

-   Identify key drivers for store profitability and shrink

-   Optimize butcher block lengths based on performance.

## Approach

The project uses gradient boosting models (XGBoost) for classifying store performance. Rather than using relative rankings, we apply fixed thresholds that provide consistent, actionable insights across different stores and time periods. The analysis covers the full process from exploring the data and building features to developing models and creating practical recommendations. By converting annual data to weekly metrics and setting clear performance benchmarks, we create tools that Albertsons can use to improve their meat and seafood operations.

# Setup and Configuration

```{r setup}
#| include: false

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
```

```{r libraries}
library(tidyverse)    
library(tidymodels)
library(vip)
library(xgboost)
library(patchwork)
library(knitr)
library(scales)
library(gridExtra)
library(dplyr)
library(ggplot2)
library(scales)
library(patchwork)


theme_set(theme_minimal(base_size = 12))
set.seed(2024)

store_colors <- c(
  "High" = "#2E7D32",
  "Medium" = "#1976D2", 
  "Unprofitable" = "#C62828"
)
```

# Data Import and Initial Processing

```{r data-import}
data <- read_csv("C:/Users/ssin307/OneDrive - Safeway, Inc/Desktop/Final.csv") 

summarize_data <- function(df, name) {
  cat(name, "Department Summary:\n")
  cat("- Total stores:", nrow(df), "\n")
  cat("- Columns:", ncol(df), "\n")
  cat("- Memory usage:", format(object.size(df), units = "MB"), "\n\n")
}

meat_data <- data |>
  select(
    # Identifiers
    `Store ID`, Division, Banner, `Street Address`, City, ST, ZIP, `District #`, Urbanicity,
    # Demographics
    `<%$45k`, `%$45-100k`, `>%$100k`, `%Hispanic/Latino`, `%White`, 
    `%Black/African American`, `%Asian`, `%Other or Multiple`,
    # Sales metrics
    `Meat Sales`, `Total Sales`, 
    # Shrink metrics
    `Meat Shrink`, `Meat Dept Shrink Pct`,
    # Operational metrics
    `Meat Markdown`, `Meat Markdown Quantity`, `Weekly Labor Hours`,
    # Butcher block metrics
    `Meat Butcher Block`
  )

seafood_data <- data |>
  select(
    # Identifiers
    `Store ID`, Division, Banner, `Street Address`, City, ST, ZIP, `District #`, Urbanicity,
    # Demographics
    `<%$45k`, `%$45-100k`, `>%$100k`, `%Hispanic/Latino`, `%White`, 
    `%Black/African American`, `%Asian`, `%Other or Multiple`,
    # Sales metrics
    `Seafood Sales`, `Total Sales`,
    # Shrink metrics
    `Seafood Shrink`, `Seafood Dept Shrink Pct`,
    # Operational metrics 
    `Seafood Markdown`, `Seafood Markdown Quantity`, `Weekly Labor Hours`,
    # Butcher block metrics
    `Seafood Butcher Block`
  )

```

# Data Cleaning and Feature Engineering

## Convert to Weekly Metrics

```{r}
character_cols <- c("Store ID", "Division", "Banner", "Street Address", "City", 
                   "ST", "ZIP", "District #", "Urbanicity")

data <- data|>
  mutate(across(
    .cols = -any_of(character_cols),
    .fns = ~ as.numeric(.)
  ))

meat_data <- meat_data|>
  mutate(across(
    .cols = -any_of(character_cols),
    .fns = ~ as.numeric(.)
  ))


seafood_data <- seafood_data|>
  mutate(across(
    .cols = -any_of(character_cols),
    .fns = ~ as.numeric(.)
  )) 




```

## Create Performance Categories

```{r}

meat_data <- meat_data |>
  mutate(performance_category = case_when(
    `Meat Sales` > 3000 & `Meat Dept Shrink Pct` < 0.04 & `Weekly Labor Hours` > 84 ~ "High",
    `Meat Sales` > 1000 & `Meat Dept Shrink Pct` < 0.05 & `Weekly Labor Hours` >= 28 ~ "Medium",
    TRUE ~ "Unprofitable"
  ))

seafood_data <- seafood_data |>
  mutate(performance_category = case_when(
    `Seafood Sales` > 3000 & `Seafood Dept Shrink Pct` < 0.04 & `Weekly Labor Hours` > 84 ~ "High",
    `Seafood Sales` > 1000 & `Seafood Dept Shrink Pct` < 0.05 & `Weekly Labor Hours` >= 28 ~ "Medium",
    TRUE ~ "Unprofitable"
  ))

meat_performance_category <- meat_data$performance_category
seafood_performance_category <- seafood_data$performance_category

performance_categories <- meat_data |>
  select(`Store ID`) |>
  mutate(
    Meat_Performance_Category = meat_performance_category,
    Seafood_Performance_Category = seafood_performance_category
  )

cat("\nMeat Performance Distribution:\n")
print(table(meat_performance_category))

cat("\nSeafood Performance Distribution:\n")
print(table(seafood_performance_category))

```

\

# Exploratory Data Analysis

## Performance Category Visualization

The meat department performance is segmented into three categories: High, Medium, and Unprofitable, based on sales, shrink percentage, and labor hours. They are categorized based on:

-   **Meat**

    1.  High performance: \>\$3,000 in Meat Sales, \<4% in shrink pct, and ? 84 weekly labor hours

    2.  Medium performance: \>\$1,000 in Meat Sales, \>4% and \<5% in Shrink pct, and \>28 labor hours per week

    3.  Any store that does not meet these criteria is labeled Unprofitable

-   **Seafood**

    1.  High performance: \>\$3,000 in Seafood Sales, \<4% in shrink pct, and ? 84 weekly labor hours

    2.  Medium performance: \>\$1,000 in Seafood Sales, \>4% and \<5% in Shrink pct, and \>28 labor hours per week

    3.  Any store that does not meet these criteria is labeled Unprofitable

```{r}
meat_counts <- meat_data |>
  count(performance_category) |>
  mutate(
    pct = n / sum(n),
    label = paste0(n, "\n", percent(pct, 0.1))
  )

p1 <- ggplot(meat_counts, aes(x = performance_category, y = n, fill = performance_category)) +
  geom_col(alpha = 0.9) +
  geom_text(aes(label = label), vjust = -0.1, color = "black", 
            fontface = "bold", size = 4) +
  scale_fill_manual(values = store_colors) +
  scale_y_continuous(limits  = c(0, max(meat_counts$n) * 2)) +
  labs(
    title = "Meat Department Performance Distribution",
    x = NULL,
    y = "Number of Stores"
  ) +
  theme(legend.position = "none")

seafood_counts <- seafood_data |>
  count(performance_category) |>
  mutate(
    pct = n / sum(n),
    label = paste0(n, "\n", percent(pct, 0.1))
  )

p2 <- ggplot(seafood_counts, aes(x = performance_category, y = n, fill = performance_category)) +
  geom_col(alpha = 0.9) +
  geom_text(aes(label = label), vjust = -0.1, color = "black", 
            fontface = "bold", size = 4) +
  scale_fill_manual(values = store_colors) +
  scale_y_continuous(limits  = c(0, max(seafood_counts$n) * 2)) +
  labs(
    title = "Seafood Department Performance Distribution",
    x = NULL,
    y = "Number of Stores"
  ) +
  theme(legend.position = "none")

p1 / p2

```

## Key Metrics by Performance Tier

### Meat Department

```{r metrics-by-tier}
#| fig-height: 10
p1 <- ggplot(meat_data, aes(x = performance_category, y = `Meat Sales`, 
                     fill = performance_category)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = store_colors) +
  scale_y_continuous(breaks  = seq(0,14000, by = 1000), labels = seq(0,14000, by = 1000)) +
  labs(x = NULL, y = "Weekly Sales", title = "Sales Distribution") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45,hjust = 1))


p3 <- ggplot(meat_data, aes(x = performance_category, y = `Meat Dept Shrink Pct`, 
                     fill = performance_category)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = store_colors) +
  scale_y_continuous(breaks  = seq(-0.5, 0.5, by = 0.1), labels = scales::percent) + 
  labs(x = NULL, y = "Shrink %", title = "Shrink Distribution") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))



p5 <- ggplot(meat_data, aes(x = performance_category, y = `Weekly Labor Hours`, 
                     fill = performance_category)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = store_colors) +
  scale_y_continuous(breaks  = seq(0, 250, by = 25), labels = seq(0, 250, by = 25)) + 
  labs(x = NULL, y = "Labor Hours", title = "Labor Distribution") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))



(p1 | p3 | p5)

```

### Seafood Department

```{r metrics-by-tier2}
#| fig-height: 10

p2 <- ggplot(seafood_data, aes(x = performance_category, y = `Seafood Sales`, 
                     fill = performance_category)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = store_colors) +
  scale_y_continuous(breaks  = seq(0,14000, by = 1000), labels = seq(0,14000, by = 1000))  +
  labs(x = NULL, y = "Weekly Sales", title = "Sales Distribution") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))


p4 <- ggplot(seafood_data, aes(x = performance_category, y =  `Seafood Dept Shrink Pct`, 
                     fill = performance_category)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = store_colors) +
  scale_y_continuous(breaks  = seq(-0.5, 0.5, by = 0.1), labels = scales::percent) + 
  labs(x = NULL, y = "Shrink %", title = "Shrink Distribution") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))


p6 <- ggplot(seafood_data, aes(x = performance_category, y = `Weekly Labor Hours`, 
                     fill = performance_category)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = store_colors) +
  scale_y_continuous(breaks  = seq(0, 250, by = 25), labels = seq(0, 250, by = 25)) + 
  labs(x = NULL, y = "Labor Hours", title = "Labor Distribution") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))


(p2| p4| p6)

```

## Performance Scatter Analysis

```{r scatter-analysis}
#| fig-height: 8
create_performance_scatter <- function(df, sales_col, shrink_col, dept) {
  # Calculate sales per labor hour
  df <- df |>
    mutate(Sales_per_Hour = get(sales_col) / `Weekly Labor Hours`)
  
  p1 <- ggplot(df, aes(x = get(sales_col), y = get(shrink_col), 
                       color = performance_category)) +
    geom_point(aes(size = `Weekly Labor Hours`), alpha = 0.7) +
    scale_color_manual(values = store_colors) +
    scale_x_continuous(labels = dollar) +
    scale_y_continuous(labels = percent) +
    scale_size_continuous(range = c(2, 8)) +
    labs(
      x = "Weekly Sales",
      y = "Shrink %",
      title = paste(dept, "Department: Sales vs Shrink"),
      subtitle = "Bubble size represents labor hours",
      color = "Performance Tier",
      size = "Labor Hours"
    ) +
    theme(legend.position = "right")
  
  p2 <- ggplot(df, aes(x = Sales_per_Hour, y = get(shrink_col), 
                       color = performance_category)) +
    geom_point(size = 3, alpha = 0.7) +
    scale_color_manual(values = store_colors) +
    scale_x_continuous(labels = dollar) +
    scale_y_continuous(labels = percent) +
    labs(
      x = "Sales per Labor Hour",
      y = "Shrink %",
      title = paste(dept, "Department: Efficiency vs Shrink"),
      color = "Performance Tier"
    ) +
    theme(legend.position = "right")
  
  p1 / p2
}

create_performance_scatter(meat_data, "Meat Sales", "Meat Dept Shrink Pct", "Meat")
```

# Classification Model

## Performance Classification Model

```{r classification-model}

class_modeling_data <- meat_data |>
  select(performance_category, `Meat Sales`, `Meat Dept Shrink Pct`, `Meat Markdown`, 
         `Meat Markdown Quantity`, `Weekly Labor Hours`, `Meat Butcher Block`,
         Division, `District #`, Urbanicity, `<%$45k`, `%$45-100k`, `>%$100k`
  ) |>
  mutate(performance_category = as.factor(performance_category))

# Create Splits for test/train
set.seed(2024)
class_split <- initial_split(class_modeling_data, prop = 0.75, 
                            strata = performance_category)
class_train <- training(class_split)
class_test <- testing(class_split)

class_recipe <- recipe(performance_category ~ ., data = class_train) |>
  step_novel(all_nominal_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())

xgb_class_spec <- boost_tree(
  trees = 1000,
  tree_depth = 6,
  min_n = 10,
  loss_reduction = 0.01,
  sample_size = 0.8,
  mtry = 1,
  learn_rate = 0.01
) |>
  set_engine("xgboost", nthread = 4) |>
  set_mode("classification")

class_workflow <- workflow() |>
  add_recipe(class_recipe) |>
  add_model(xgb_class_spec)

class_fit <- class_workflow |>
  fit(data = class_train)

class_predictions <- predict(class_fit, class_test) |>
  bind_cols(predict(class_fit, class_test, type = "prob")) |>
  bind_cols(class_test)

metrics <- metric_set(accuracy, recall)
class_metrics <- class_predictions |>
  metrics(truth = performance_category, estimate = .pred_class)

conf_mat <- class_predictions |>
  conf_mat(truth = performance_category, estimate = .pred_class)

class_metrics |>
  filter(.metric %in% c("accuracy", "kap")) |>
  mutate(.estimate = round(.estimate, 3)) |>
  kable(col.names = c("Metric", "Estimator", "Value"))

cat("\nPrediction VS Truth:\n")
print(conf_mat)
```

## Feature Importance Analysis

```{r feature-importance}
#| fig-height: 10

# Extract feature importance for classification model
class_vip <- class_fit |>
  extract_fit_parsnip() |>
  vip(num_features = 10) +
  labs(title = "Top 10 Features for Performance Classification") +
  theme_minimal()

class_vip
```

# Store Segmentation Analysis

### Meat butcher Block

```{r segmentation meat}
# Create detailed segments
segmentation_data <- meat_data |>
  mutate(
    Sales_per_Hour = `Meat Sales` / `Weekly Labor Hours`,
    Markdown_Rate = `Meat Markdown` / `Meat Sales`,
    Sales_per_Foot = `Meat Sales` / `Meat Butcher Block`,
    
    Revenue_Segment = cut(
      `Meat Sales`,
      breaks = quantile(`Meat Sales`, probs = c(0, 0.33, 0.67, 1), na.rm = TRUE),
      labels = c("Low Revenue", "Mid Revenue", "High Revenue"),
      include.lowest = TRUE
    ),
    
    Efficiency_Segment = cut(
      Sales_per_Hour,
      breaks = quantile(Sales_per_Hour, probs = c(0, 0.33, 0.67, 1), na.rm = TRUE),
      labels = c("Low Efficiency", "Mid Efficiency", "High Efficiency"),
      include.lowest = TRUE
    ),
    
    Shrink_Control = cut(
      `Meat Dept Shrink Pct`,
      breaks = quantile(`Meat Dept Shrink Pct`, probs = c(0, 0.33, 0.67, 1), na.rm = TRUE),
      labels = c("Poor Control", "Average Control", "Good Control"),
      include.lowest = TRUE
    )
  )

segment_summary <- segmentation_data |>
  group_by(Revenue_Segment, Shrink_Control) |>
  summarise(
    Store_Count = n(),
    Avg_Sales = mean(`Meat Sales`),
    Avg_Shrink_Pct = mean(`Meat Dept Shrink Pct`),
    Avg_Efficiency = mean(Sales_per_Hour),
    Performance_Distribution = list(table(performance_category)),
    .groups = "drop"
  ) |>
  arrange(desc(Avg_Sales))

segment_summary |>
  select(-Performance_Distribution) |>
  mutate(
    Avg_Sales = dollar(Avg_Sales),
    Avg_Shrink_Pct = percent(Avg_Shrink_Pct, 0.01),
    Avg_Efficiency = dollar(Avg_Efficiency)
  ) |>
  kable(
    col.names = c("Revenue", "Shrink Control", "Stores", "Avg Sales", 
                  "Avg Shrink", "Sales/Hour"),
    align = c("l", "l", "r", "r", "r", "r")
  )
```

### Seafood Butcher Block

```{r segmentation seafood}
# Create detailed segments
segmentation_data <- seafood_data |>
  mutate(
    Sales_per_Hour = `Seafood Sales` / `Weekly Labor Hours`,
    Markdown_Rate = `Seafood Markdown` / `Seafood Sales`,
    Sales_per_Foot = `Seafood Sales` / `Seafood Butcher Block`,
    
    Revenue_Segment = cut(
      `Seafood Sales`,
      breaks = quantile(`Seafood Sales`, probs = c(0, 0.33, 0.67, 1), na.rm = TRUE),
      labels = c("Low Revenue", "Mid Revenue", "High Revenue"),
      include.lowest = TRUE
    ),
    
    Efficiency_Segment = cut(
      Sales_per_Hour,
      breaks = quantile(Sales_per_Hour, probs = c(0, 0.33, 0.67, 1), na.rm = TRUE),
      labels = c("Low Efficiency", "Mid Efficiency", "High Efficiency"),
      include.lowest = TRUE
    ),
    
    Shrink_Control = cut(
      `Seafood Dept Shrink Pct`,
      breaks = quantile(`Seafood Dept Shrink Pct`, probs = c(0, 0.33, 0.67, 1), na.rm = TRUE),
      labels = c("Poor Control", "Average Control", "Good Control"),
      include.lowest = TRUE
    )
  )

segment_summary <- segmentation_data |>
  group_by(Revenue_Segment, Shrink_Control) |>
  summarise(
    Store_Count = n(),
    Avg_Sales = mean(`Seafood Sales`),
    Avg_Shrink_Pct = mean(`Seafood Dept Shrink Pct`),
    Avg_Efficiency = mean(Sales_per_Hour),
    Performance_Distribution = list(table(performance_category)),
    .groups = "drop"
  ) |>
  arrange(desc(Avg_Sales))

segment_summary |>
  select(-Performance_Distribution) |>
  mutate(
    Avg_Sales = dollar(Avg_Sales),
    Avg_Shrink_Pct = percent(Avg_Shrink_Pct, 0.01),
    Avg_Efficiency = dollar(Avg_Efficiency)
  ) |>
  kable(
    col.names = c("Revenue", "Shrink Control", "Stores", "Avg Sales", 
                  "Avg Shrink", "Sales/Hour"),
    align = c("l", "l", "r", "r", "r", "r")
  )
```

# Shrink Prediction Model

```{r shrink-analysis-model}
# Create shrink analysis dataset
shrink_modeling_data <- meat_data |>
  select(
    `Store ID`,
    # Target variable
    `Meat Shrink`,
    # Operational features
    `Meat Sales`, `Meat Markdown`, `Meat Markdown Quantity`, 
    `Weekly Labor Hours`, `Meat Butcher Block`,
    # Store characteristics
    Division, `District #`, Urbanicity,
    # Demographics
    `<%$45k`, `%$45-100k`, `>%$100k`, 
    `%Hispanic/Latino`, `%White`, `%Black/African American`, 
    `%Asian`, `%Other or Multiple`, `performance_category`
  ) |>
  mutate(performance_category = as.factor(performance_category))
shrink_modeling_data[is.na(shrink_modeling_data)] <- 0

# Create splits for shrink prediction
set.seed(2024)
shrink_split <- initial_split(shrink_modeling_data, prop = 0.75)
shrink_train <- training(shrink_split)
shrink_test <- testing(shrink_split)
shrink_train
# Recipe for shrink prediction
shrink_recipe <- recipe(`Meat Shrink` ~ ., data = shrink_train) |>
  step_novel(all_nominal_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())

# XGBoost model for shrink prediction (regression)
xgb_shrink_spec <- boost_tree(
  trees = 1000,
  tree_depth = tune(),
  min_n = tune(),
  loss_reduction = tune(),
  sample_size = tune(),
  mtry = tune(),
  learn_rate = 0.01
) |>
  set_engine("xgboost", nthread = 4) |>
  set_mode("regression")

shrink_workflow <- workflow() |>
  add_recipe(shrink_recipe) |>
  add_model(xgb_shrink_spec)

xgb_grid <- grid_latin_hypercube(
  tree_depth(range = c(3, 10)),
  min_n(range = c(5, 50)),
  loss_reduction(range = c(-10,1.5), trans = log10_trans()),
  sample_size(range = c(0,1)),
  mtry(range = c(3,7)),
  size = 30
)

shrink_folds <- vfold_cv(shrink_train, v = 2, repeats = 2, strata = Division)

# Tune model
doParallel::registerDoParallel(cores = 4)

shrink_tune_results <- tune_grid(
  shrink_workflow,
  resamples = shrink_folds,
  grid = xgb_grid,
  metrics = metric_set(rmse,rsq, mae), 
  control = control_grid(verbose = FALSE, save_pred = TRUE)
)

best_shrink <- select_best(shrink_tune_results, metric = "rmse")

shrink_workflow <- finalize_workflow(shrink_workflow, best_shrink)

shrink_fit <- shrink_workflow |>
  fit(data = shrink_train)

# Make predictions
shrink_predictions <- predict(shrink_fit, shrink_test) |>
  bind_cols(shrink_test)


metrics <- metric_set(rmse, rsq, mae)

shrink_metrics <- shrink_predictions |>
  metrics(truth = `Meat Shrink`, estimate = .pred)

shrink_metrics |>
  mutate(
    .estimate = case_when(
      .metric == "rmse" ~ dollar(.estimate),
      .metric == "mae" ~ dollar(.estimate),
      TRUE ~ as.character(round(.estimate, 3))
    )
  ) |>
  kable(col.names = c("Metric", "Estimator", "Value"))


# View actual vs predicted shrink values for the test set
shrink_predictions |> 
  select(`Store ID`, `Meat Shrink`, .pred, performance_category) 
ggplot(shrink_predictions, aes(x = `Meat Shrink`, y = .pred)) +
  geom_point(alpha = 0.6) +
  geom_abline(linetype = "dashed", color = "red") +
  labs(
    title = "Actual vs Predicted",
    x = "Actual Shrink",
    y = "Predicted Shrink"
  ) +
  theme_minimal()

```

## Feature Importance for Shrink

```{r shrink-feature-importance}
#| fig-height: 10

# Extract feature importance for shrink model
shrink_vip <- shrink_fit |>
  extract_fit_parsnip() |>
  vip(num_features = 10) +
  labs(
    title = "Top 10 Features Influencing Meat Department Shrink",
    subtitle = "Higher importance indicates stronger influence on shrink rates"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

shrink_vip

# Get the actual importance values for interpretation
shrink_importance <- shrink_fit |>
  extract_fit_parsnip() |>
  vip::vi() |>
  slice_head(n = 10) |>
  mutate(
    Importance_Pct = (Importance / sum(Importance)) * 100,
    Interpretation = case_when(
      Variable == "Meat_Sales" ~ "Higher sales correlate with shrink patterns",
      Variable == "Weekly_Labor_Hours" ~ "Labor allocation affects shrink control",
      Variable == "Meat_Markdown" ~ "Markdown practices influence shrink",
      Variable == "Meat_Markdown_Quantity" ~ "Volume of markdowns impacts shrink",
      Variable == "Meat_Butcher_Block" ~ "Display size affects product turnover",
      grepl("Division", Variable) ~ "Regional differences in shrink management",
      grepl("District", Variable) ~ "District-level operational differences",
      grepl("Urbanicity", Variable) ~ "Urban vs rural shrink patterns",
      grepl("%", Variable) ~ "Demographics influence shopping patterns",
      TRUE ~ "Store characteristic affecting shrink"
    )
  )

cat("\n=== SHRINK FACTOR IMPORTANCE BREAKDOWN ===\n")
shrink_importance |>
  select(Variable, Importance, Importance_Pct, Interpretation) |>
  mutate(
    Importance = round(Importance, 3),
    Importance_Pct = paste0(round(Importance_Pct, 1), "%")
  ) |>
  kable(
    col.names = c("Feature", "Importance Score", "% of Total", "Interpretation"),
    caption = "Detailed breakdown of factors influencing meat department shrink"
  )
```

## Combined Feature Importance Visualization

```{r combined-vip-plots}
#| fig-height: 12

# Combine both VIP plots for comparison
combined_vip <- (class_vip / shrink_vip) +
  plot_annotation(
    title = "Feature Importance Analysis: Performance Classification vs Shrink Prediction",
    subtitle = "Comparing key drivers of store performance tiers and shrink rates",
    theme = theme(
      plot.title = element_text(face = "bold", size = 16),
      plot.subtitle = element_text(size = 14)
    )
  )

combined_vip
```

## Shrink Analysis by Performance Category

```{r shrink-by-performance}
#| fig-height: 8

# Analyze shrink patterns by performance category
shrink_performance_analysis <- meat_data |>
  group_by(performance_category) |>
  summarise(
    Store_Count = n(),
    Avg_Shrink_Pct = mean(`Meat Dept Shrink Pct`, na.rm = TRUE),
    Median_Shrink_Pct = median(`Meat Dept Shrink Pct`, na.rm = TRUE),
    Min_Shrink_Pct = min(`Meat Dept Shrink Pct`, na.rm = TRUE),
    Max_Shrink_Pct = max(`Meat Dept Shrink Pct`, na.rm = TRUE),
    SD_Shrink_Pct = sd(`Meat Dept Shrink Pct`, na.rm = TRUE),
    .groups = "drop"
  )


shrink_performance_analysis |>
  mutate(
    Avg_Shrink_Pct = percent(Avg_Shrink_Pct, 0.01),
    Median_Shrink_Pct = percent(Median_Shrink_Pct, 0.01),
    Min_Shrink_Pct = percent(Min_Shrink_Pct, 0.01),
    Max_Shrink_Pct = percent(Max_Shrink_Pct, 0.01),
    SD_Shrink_Pct = percent(SD_Shrink_Pct, 0.01)
  ) |>
  kable(
    col.names = c("Performance", "Stores", "Avg Shrink", "Median Shrink", 
                  "Max Shrink", "Min Shrink", "Std Dev"),
    caption = "Shrink statistics by store performance category"
  )

# Visualization of shrink distribution by performance
p_shrink_dist <- ggplot(meat_data, aes(x = performance_category, y = `Meat Dept Shrink Pct`, 
                                       fill = performance_category)) +
  geom_violin(alpha = 0.7) +
  geom_boxplot(width = 0.2, alpha = 0.9) +
  scale_fill_manual(values = store_colors) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Shrink Distribution by Performance Category",
    subtitle = "Violin plots show distribution density, box plots show quartiles",
    x = "Performance Category",
    y = "Shrink Percentage"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Scatter plot showing relationship between key factors and shrink
p_shrink_factors <- meat_data |>
  mutate(Sales_per_Labor_Hour = `Meat Sales` / `Weekly Labor Hours`) |>
  ggplot(aes(x = Sales_per_Labor_Hour, y = `Meat Dept Shrink Pct`, 
             color = performance_category)) +
  geom_point(aes(size = `Meat Markdown`), alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed") +
  scale_color_manual(values = store_colors) +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(labels = dollar) +
  scale_size_continuous(labels = dollar, range = c(2, 8)) +
  labs(
    title = "Sales Efficiency vs Shrink Rate",
    subtitle = "Bubble size represents markdown amount",
    x = "Sales per Labor Hour",
    y = "Shrink Percentage",
    color = "Performance",
    size = "Markdown $"
  ) +
  theme_minimal()

p_shrink_dist / p_shrink_factors
```

# Butcher Block Setup

```{r butcher-block-setup}
#Define butcher block constraints and parameters
MEAT_BASELINE_LENGTH <- 8 
SEAFOOD_BASELINE_LENGTH <- 12
MIN_MEAT_LENGTH <- 4       
MIN_SEAFOOD_LENGTH <- 8      
MAX_MEAT_LENGTH <- 50        
MAX_SEAFOOD_LENGTH <- 50     
ADJUSTMENT_INCREMENT <- 4    # adjust in 4-foot increments

#Operational expense parameters
OPERATIONAL_EXPENSE_PER_BLOCK <- 1500  # $1500 per 4ft block
```

## Data Preparations

```{r butcher-block-data-prep}

butcher_data <- data |>
  select(`Store ID`, Division, Banner, `Street Address`, City, ST, ZIP, `District #`, Urbanicity,`<%$45k`, `%$45-100k`, `>%$100k`, `%Hispanic/Latino`, `%White`, `%Black/African American`, `%Asian`, `%Other or Multiple`,`Meat Sales`, `Seafood Sales`, `Total Sales`,`Meat Shrink`, `Seafood Shrink`, `Meat Dept Shrink Pct`, `Seafood Dept Shrink Pct`,`Meat Butcher Block`, `Seafood Butcher Block`, `Total Butcher Block Length`,`Meat Markdown`, `Seafood Markdown`, `Meat Markdown Quantity`, `Seafood Markdown Quantity`,`Total Markdown`, `Total Markdown Quantity`, `Weekly Labor Hours`
  ) |>
  mutate(across(
    .cols = -any_of(character_cols),
    .fns = ~ as.numeric(.)
  )) |>
  mutate(
    `Meat Butcher Block` = as.numeric(`Meat Butcher Block`),
    `Seafood Butcher Block` = as.numeric(`Seafood Butcher Block`)
  )|>
  mutate(
    Meat_Sales_per_Foot = `Meat Sales` / `Meat Butcher Block`,
    Seafood_Sales_per_Foot = `Seafood Sales` / `Seafood Butcher Block`,
    
    Meat_Shrink_per_Foot = `Meat Shrink` / `Meat Butcher Block`,
    Seafood_Shrink_per_Foot = `Seafood Shrink` / `Seafood Butcher Block`,
    
    Meat_Length_Efficiency = Meat_Sales_per_Foot / (MEAT_BASELINE_LENGTH * 100),
    Seafood_Length_Efficiency = Seafood_Sales_per_Foot / (SEAFOOD_BASELINE_LENGTH * 100)
  )

butcher_data <- butcher_data |>
  mutate(
    meat_performance = case_when(
      `Meat Sales` > 3000 & `Meat Dept Shrink Pct` < 0.04 & `Weekly Labor Hours` > 84 ~ "High",
      `Meat Sales` > 1000 & `Meat Dept Shrink Pct` < 0.05 & `Weekly Labor Hours` >= 28 ~ "Medium",
      TRUE ~ "Unprofitable"
    ),
    seafood_performance = case_when(
      `Seafood Sales` > 3000 & `Seafood Dept Shrink Pct` < 0.04 & `Weekly Labor Hours` > 84 ~ "High",
      `Seafood Sales` > 1000 & `Seafood Dept Shrink Pct` < 0.05 & `Weekly Labor Hours` >= 28 ~ "Medium",
      TRUE ~ "Unprofitable"
    )
  )

cat("Butcher Block Analysis Dataset prepared with", nrow(butcher_data), "stores\n")

```

## Exploratory Data Analysis

```{r butcher-block-eda}

p1 <- ggplot(butcher_data, aes(x = `Meat Butcher Block`, y = Meat_Sales_per_Foot, 
                               color = meat_performance)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_vline(xintercept = MEAT_BASELINE_LENGTH, linetype = "dashed", color = "red") +
  scale_color_manual(values = store_colors) +
  scale_y_continuous(labels = dollar) +
  labs(
    title = "Meat Department: Block Length vs Sales Efficiency",
    x = "Butcher Block Length (feet)",
    y = "Sales per Foot",
    color = "Performance",
    subtitle = "Red line shows baseline length (8 ft)"
  )

p2 <- ggplot(butcher_data, aes(x = `Seafood Butcher Block`, y = Seafood_Sales_per_Foot,
                               color = seafood_performance)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_vline(xintercept = SEAFOOD_BASELINE_LENGTH, linetype = "dashed", color = "red") +
  scale_color_manual(values = store_colors) +
  scale_y_continuous(labels = dollar) +
  labs(
    title = "Seafood Department: Block Length vs Sales Efficiency",
    x = "Butcher Block Length (feet)",
    y = "Sales per Foot",
    color = "Performance",
    subtitle = "Red line shows baseline length (12 ft)"
  )

grid.arrange(p1, p2, ncol = 1)
```

## Modelling

```{r optimization-model}
create_butcher_block_recommendations <- function(data) {
  recommendations <- data |>
    mutate(
      #Define sales thresholds based on your performance categories
      sales_threshold_meat = 1000,    # Weekly sales threshold for meat
      sales_threshold_seafood = 1000, # Weekly sales threshold for seafood
      
      Meat_Performance_Pct = (`Meat Sales` - sales_threshold_meat) / sales_threshold_meat * 100,
      Seafood_Performance_Pct = (`Seafood Sales` - sales_threshold_seafood) / sales_threshold_seafood * 100,
      Meat_Raw_Adjustment = case_when(
        Meat_Performance_Pct <= -70 & `Meat Butcher Block` > 11 ~ -8,
        Meat_Performance_Pct <= -30 & `Meat Butcher Block` > 11 ~ -4,
        TRUE ~ 0
      ),
      
      Seafood_Raw_Adjustment = case_when(
        Seafood_Performance_Pct <= -70 & `Seafood Butcher Block` > 15 ~ -8,
        Seafood_Performance_Pct <= -30 & `Seafood Butcher Block` > 15 ~ -4,
        TRUE ~ 0
      ),
      Meat_Optimal_Length = case_when(
        (`Meat Butcher Block` + Meat_Raw_Adjustment) < MIN_MEAT_LENGTH ~ 
          pmax(MIN_MEAT_LENGTH, `Meat Butcher Block` - 4),
        Meat_Raw_Adjustment < 0 ~ 
          `Meat Butcher Block` + (floor(abs(Meat_Raw_Adjustment) / 4) * -4),
        Meat_Raw_Adjustment > 0 ~ 
          pmin(`Meat Butcher Block` + (ceiling(Meat_Raw_Adjustment / 4) * 4), MAX_MEAT_LENGTH),
        TRUE ~ `Meat Butcher Block`
      ),
      Seafood_Optimal_Length = case_when(
        (`Seafood Butcher Block` + Seafood_Raw_Adjustment) < MIN_SEAFOOD_LENGTH ~ 
          pmax(MIN_SEAFOOD_LENGTH, `Seafood Butcher Block` - 4),
        Seafood_Raw_Adjustment < 0 ~ 
          `Seafood Butcher Block` + (floor(abs(Seafood_Raw_Adjustment) / 4) * -4),
        Seafood_Raw_Adjustment > 0 ~ 
          pmin(`Seafood Butcher Block` + (ceiling(Seafood_Raw_Adjustment / 4) * 4), MAX_SEAFOOD_LENGTH),
        TRUE ~ `Seafood Butcher Block`
      )
    ) |>
    mutate(
      # Calculate actual changes
      Meat_Length_Change = Meat_Optimal_Length - `Meat Butcher Block`,
      Seafood_Length_Change = Seafood_Optimal_Length - `Seafood Butcher Block`,
      
      Meat_Priority = case_when(
        Meat_Performance_Pct <= -70 & `Meat Butcher Block` > 12 & Meat_Length_Change == -8 ~ "HIGH",
        Meat_Performance_Pct <= -40 & `Meat Butcher Block` > 12 & Meat_Length_Change == -4 ~ "HIGH",
        Meat_Performance_Pct <= -40 & `Meat Butcher Block` <= 12 ~ "MEDIUM",
        TRUE ~ "LOW"
      ),
      
      Seafood_Priority = case_when(
        Seafood_Performance_Pct <= -70 & `Seafood Butcher Block` > 16 & Seafood_Length_Change == -8 ~ "HIGH",
        Seafood_Performance_Pct <= -40 & `Seafood Butcher Block` > 16 & Seafood_Length_Change == -4 ~ "HIGH", 
        Seafood_Performance_Pct <= -40 & `Seafood Butcher Block` <= 16 ~ "MEDIUM",
        TRUE ~ "LOW"
      ),
      
      Meat_Recommendation = case_when(
        Meat_Length_Change == -8 ~ 
          paste("DECREASE by 8 feet (2 blocks) - Performance", round(abs(Meat_Performance_Pct), 1), "% below threshold, store has >12ft"),
        Meat_Length_Change == -4 ~ 
          paste("DECREASE by 4 feet (1 block) - Performance", round(abs(Meat_Performance_Pct), 1), "% below threshold, store has >12ft"),
        Meat_Performance_Pct <= -40 & `Meat Butcher Block` <= 12 ~
          paste("MONITOR - Performance", round(abs(Meat_Performance_Pct), 1), "% below threshold, but store only has", `Meat Butcher Block`, "ft (≤12ft baseline)"),
        TRUE ~ "MAINTAIN current length"
      ),
      
      Seafood_Recommendation = case_when(
        Seafood_Length_Change == -8 ~ 
          paste("DECREASE by 8 feet (2 blocks) - Performance", round(abs(Seafood_Performance_Pct), 1), "% below threshold, store has >16ft"),
        Seafood_Length_Change == -4 ~ 
          paste("DECREASE by 4 feet (1 block) - Performance", round(abs(Seafood_Performance_Pct), 1), "% below threshold, store has >16ft"),
        Seafood_Performance_Pct <= -40 & `Seafood Butcher Block` <= 16 ~
          paste("MONITOR - Performance", round(abs(Seafood_Performance_Pct), 1), "% below threshold, but store only has", `Seafood Butcher Block`, "ft (≤16ft baseline)"),
        TRUE ~ "MAINTAIN current length"
      ),
      
      # Calculate operational expense saved
      Meat_Operational_Expense_Saved = case_when(
        Meat_Length_Change < 0 ~ abs(Meat_Length_Change) / 4 * OPERATIONAL_EXPENSE_PER_BLOCK,
        TRUE ~ 0
      ),
      Seafood_Operational_Expense_Saved = case_when(
        Seafood_Length_Change < 0 ~ abs(Seafood_Length_Change) / 4 * OPERATIONAL_EXPENSE_PER_BLOCK,
        TRUE ~ 0
      ),
      Total_Operational_Expense = Meat_Operational_Expense_Saved + Seafood_Operational_Expense_Saved,
      
      Estimated_Meat_Sales_Impact = case_when(
        Meat_Length_Change < 0 ~ Meat_Length_Change * Meat_Sales_per_Foot * 0.9, # Sales loss from reduction
        TRUE ~ 0
      ),
      
      Estimated_Seafood_Sales_Impact = case_when(
        Seafood_Length_Change < 0 ~ Seafood_Length_Change * Seafood_Sales_per_Foot * 0.9,
        TRUE ~ 0
      ),
      
      Total_Sales_Impact = Estimated_Meat_Sales_Impact + Estimated_Seafood_Sales_Impact,
      
      # Break-even calculation in months based on operational expense savings vs sales loss
      Breakeven_Months = case_when(
        Total_Operational_Expense > 0 & Total_Sales_Impact < 0 ~ 
          Total_Operational_Expense / (abs(Total_Sales_Impact) * 4.33), # 4.33 weeks per month
        TRUE ~ 0
      ),

      Priority_Score = case_when(
        Meat_Priority == "HIGH" | Seafood_Priority == "HIGH" ~ "HIGH",
        Meat_Priority == "MEDIUM" | Seafood_Priority == "MEDIUM" ~ "MEDIUM", 
        TRUE ~ "NO CHANGE"
      ),

      Overall_Performance = case_when(
        (Meat_Performance_Pct >= 20 & Seafood_Performance_Pct >= 20) ~ "High Performer",
        (Meat_Performance_Pct >= -20 & Seafood_Performance_Pct >= -20) ~ "Stable Performer",
        TRUE ~ "Underperformer"
      ),

      Meat_Constraint_Check = case_when(
        Meat_Optimal_Length < MIN_MEAT_LENGTH ~ "Below minimum",
        Meat_Optimal_Length > MAX_MEAT_LENGTH ~ "Above maximum", 
        (Meat_Length_Change %% 4) != 0 ~ "Not 4ft increment",
        TRUE ~ "OK"
      ),
      
      Seafood_Constraint_Check = case_when(
        Seafood_Optimal_Length < MIN_SEAFOOD_LENGTH ~ "Below minimum",
        Seafood_Optimal_Length > MAX_SEAFOOD_LENGTH ~ "Above maximum",
        (Seafood_Length_Change %% 4) != 0 ~ "Not 4ft increment", 
        TRUE ~ "OK"
      )
    )
  
  return(recommendations)
}



##Generating recommendations

butcher_recommendations <- create_butcher_block_recommendations(butcher_data)

recommendation_summary <- butcher_recommendations |>
  summarise(
    Total_Stores = n(),
    Stores_with_Changes = sum(Meat_Length_Change != 0 | Seafood_Length_Change != 0),

    Meat_Decreases_4ft = sum(Meat_Length_Change == -4),
    Meat_Decreases_8ft = sum(Meat_Length_Change == -8), 
    Avg_Meat_Performance = round(mean(Meat_Performance_Pct, na.rm = TRUE), 1),

    Seafood_Decreases_4ft = sum(Seafood_Length_Change == -4),
    Seafood_Decreases_8ft = sum(Seafood_Length_Change == -8),
    Avg_Seafood_Performance = round(mean(Seafood_Performance_Pct, na.rm = TRUE), 1),

    High_Priority = sum(Priority_Score == "HIGH"),
    Medium_Priority = sum(Priority_Score == "MEDIUM"),
    No_Change = sum(Priority_Score == "NO CHANGE"),

    Total_Operational_Expense = round(sum(Total_Operational_Expense), 0),
    Total_Sales_Impact = round(sum(Total_Sales_Impact), 0),
    Net_Impact = round(sum(Total_Sales_Impact) + sum(Total_Operational_Expense), 0),

    Meat_Constraint_Violations = sum(Meat_Constraint_Check != "OK"),
    Seafood_Constraint_Violations = sum(Seafood_Constraint_Check != "OK"),

    High_Performers = sum(Overall_Performance == "High Performer"),
    Stable_Performers = sum(Overall_Performance == "Stable Performer"),
    Underperformers = sum(Overall_Performance == "Underperformer")
  )

constraint_issues <- butcher_recommendations |>
  filter(Meat_Constraint_Check != "OK" | Seafood_Constraint_Check != "OK") |>
  select(`Store ID`, Meat_Constraint_Check, Seafood_Constraint_Check,
         `Meat Butcher Block`, Meat_Optimal_Length,
         `Seafood Butcher Block`, Seafood_Optimal_Length)

```

## Recommendations

```{r top-recommendations}
high_impact_recommendations <- butcher_recommendations |>
  filter(Priority_Score %in% c("HIGH")) |>
  select(
    `Store ID`, Division, Banner,
    meat_performance, seafood_performance,
    `Meat Butcher Block`, Meat_Optimal_Length, Meat_Recommendation,
    `Seafood Butcher Block`, Seafood_Optimal_Length, Seafood_Recommendation,
    Total_Operational_Expense, Total_Sales_Impact, Breakeven_Months, Priority_Score
  ) |>
  arrange(desc(Priority_Score), desc(abs(Total_Operational_Expense)))

cat("\n=== BUTCHER BLOCK RECOMMENDATIONS WITH HIGH PRIORITY ===\n")
high_impact_recommendations |>
  mutate(
    Total_Operational_Expense = dollar(Total_Operational_Expense),
    Total_Sales_Impact = dollar(Total_Sales_Impact),
    Breakeven_Months = round(Breakeven_Months, 2)
  ) |>
  kable()
```

## EDA of Recommendations

```{r recommendation-visualization}
p1 <- ggplot(butcher_recommendations, aes(x = Meat_Length_Change, fill = meat_performance)) +
  geom_histogram(bins = 15, alpha = 0.7, position = "stack") +
  scale_fill_manual(values = store_colors) +
  labs(
    title = "Distribution of Meat Block Length Changes",
    x = "Recommended Length Change (feet)",
    y = "Number of Stores",
    fill = "Performance"
  ) +
  theme_minimal()

p2 <- ggplot(butcher_recommendations, aes(x = Seafood_Length_Change, fill = seafood_performance)) +
  geom_histogram(bins = 15, alpha = 0.7, position = "stack") +
  scale_fill_manual(values = store_colors) +
  labs(
    title = "Distribution of Seafood Block Length Changes",
    x = "Recommended Length Change (feet)",
    y = "Number of Stores",
    fill = "Performance"
  ) +
  theme_minimal()

p3 <- ggplot(butcher_recommendations, aes(x = Total_Operational_Expense, y = Total_Sales_Impact, 
                                         color = Priority_Score, size = abs(Breakeven_Months))) +
  geom_point(alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_x_continuous(labels = dollar) +
  scale_y_continuous(labels = dollar) +
  scale_color_manual(values = c("HIGH" = "red", "MEDIUM" = "orange", "LOW" = "yellow", "NO CHANGE" = "gray")) +
  labs(
    title = "Operational Expense Savings vs Sales Impact of Recommendations",
    x = "Operational Expense Saved",
    y = "Weekly Sales Impact",
    color = "Priority",
    size = "Breakeven Months"
  ) +
  theme_minimal()

grid.arrange(p1, p2, p3, ncol = 1)
```

## Sales impact Analysis

```{r sales-impact-analysis}
#| fig-height: 12
create_detailed_sales_impact_table <- function(recommendations_data) {
 # Moderate scenario (25% sales loss per foot reduced)
  sales_impact_analysis <- recommendations_data |>
    mutate(
     
      Meat_Loss = case_when(
        Meat_Length_Change < 0 ~ abs(Meat_Length_Change) * Meat_Sales_per_Foot * 0.25,
        TRUE ~ 0
      ),
      Seafood_Loss = case_when(
        Seafood_Length_Change < 0 ~ abs(Seafood_Length_Change) * Seafood_Sales_per_Foot * 0.25,
        TRUE ~ 0
      ),

      Weekly_Loss = Meat_Loss + Seafood_Loss,
      Annual_Loss = Weekly_Loss * 52,
      
      # Break-even calculations (months to recover operational expense through sales)
      Breakeven_Months = case_when(
        Weekly_Loss > 0 & Total_Operational_Expense > 0 ~ 
          Total_Operational_Expense / (Weekly_Loss * 4.33),
        TRUE ~ 0
      ),
      
      # Sales impact as percentage of current sales
      Sales_Impact_Pct = (Weekly_Loss / (`Meat Sales` + `Seafood Sales`)) * 100
    )
  
  return(sales_impact_analysis)
}

enhanced_recommendations <- create_detailed_sales_impact_table(butcher_recommendations)

sales_impact_summary <- enhanced_recommendations |>
  filter(Meat_Length_Change < 0 | Seafood_Length_Change < 0) |>
  select(
    `Store ID`, Division, Banner, 
    meat_performance, seafood_performance,
    `Meat Butcher Block`, `Seafood Butcher Block`,
    Meat_Length_Change, Seafood_Length_Change,
    Weekly_Loss,
    Sales_Impact_Pct,
    Breakeven_Months, Priority_Score
  ) |>
  arrange(desc(Priority_Score), desc(Weekly_Loss))


cat("=== DETAILED SALES IMPACT ANALYSIS ===\n")
cat("Shows potential weekly sales losses from butcher block reductions\n\n")
cat("Assuming moderate scenario (25% sales loss per foot reduced)\n\n")





sales_impact_simple <- sales_impact_summary |>
  head(5) |>
  select(
    `Store ID`, Banner, 
    Meat_Length_Change, Seafood_Length_Change,
    Weekly_Loss, 
    Sales_Impact_Pct, Priority_Score
  ) |>
  mutate(
    Meat_Length_Change = paste0(Meat_Length_Change, " ft"),
    Seafood_Length_Change = paste0(Seafood_Length_Change, " ft"),
    Weekly_Loss = dollar(Weekly_Loss),
    Sales_Impact_Pct = paste0(round(Sales_Impact_Pct, 1), "%")
  )

kable(sales_impact_simple, 
      col.names = c("Store ID", "Banner", "Meat Change", "Seafood Change", 
                    "Weekly Loss", "% Impact", "Priority"),
      caption = "Sales Impact Analysis - Top Stores with Reductions")

aggregate_impact <- enhanced_recommendations |>
  filter(Meat_Length_Change < 0 | Seafood_Length_Change < 0) |>
  group_by(Priority_Score) |>
  summarise(
    Store_Count = n(),
    Total_Annual_Loss = sum(Annual_Loss),
    .groups = "drop"
  ) |>
  arrange(desc(Priority_Score), desc(Total_Annual_Loss))

cat("\n=== AGGREGATE SALES IMPACT BY PRIORITY AND RISK ===\n")
aggregate_impact |>
  mutate(
    Total_Annual_Loss = dollar(Total_Annual_Loss),
  ) |>
  kable()


#Visualizations
sales_impact_viz <- enhanced_recommendations |>
  filter(Meat_Length_Change < 0 | Seafood_Length_Change < 0) |>
  select(`Store ID`, Weekly_Loss, Priority_Score) |>
  slice_head(n = 10) |>  # Limit to top 30 stores for readability
  pivot_longer(cols = c(Weekly_Loss),
               names_to = "Scenario", values_to = "Weekly_Loss") |>
  mutate(Scenario = str_remove(Scenario, "Weekly_Loss"))


total_impact_summary <- enhanced_recommendations |>
  filter(Meat_Length_Change < 0 | Seafood_Length_Change < 0) |>
  summarise(
    Total_Stores_Affected = n(),
    Company_Wide_Weekly_Loss = sum(Weekly_Loss),
  )

cat("\n=== COMPANY-WIDE SALES IMPACT SUMMARY ===\n")
cat("Total stores with recommended reductions:", total_impact_summary$Total_Stores_Affected, "\n")
cat("Company-wide weekly sales loss (moderate scenario):", dollar(total_impact_summary$Company_Wide_Weekly_Loss), "\n")

```

```{}
```

# Final Table

```{r detailed-sales-impact-table}
final <- enhanced_recommendations |>
  select(
    `Store ID`, Division, City, ST,
    `Meat Sales`, meat_performance,`Meat Butcher Block`,Meat_Length_Change, Meat_Recommendation,
    `Seafood Sales`, seafood_performance, `Seafood Butcher Block`, Seafood_Length_Change, Seafood_Recommendation,
    ,  
    Priority_Score,
    Breakeven_Months 
  ) |>
  arrange(desc(Priority_Score), desc(abs(Meat_Length_Change) + abs(Seafood_Length_Change))) |>
  mutate(
    `Meat Sales` = dollar(`Meat Sales`),
    `Seafood Sales` = dollar(`Seafood Sales`),
    Breakeven_Months = round(Breakeven_Months, 1)
  )

cat("\n=== DETAILED IMPACT TABLE FOR HIGH & MEDIUM PRIORITY STORES ===\n")
final |>
  arrange(`Store ID`) |>
  kable(
    col.names = c("Store ID", "Division", "City", "ST",
                  "Meat Sales", "meat performance",  "Meat Butcher Block", 
                  "Meat_Length_Change","Meat_Recommendation","Seafood Sales","seafood performance","Seafood Butcher Block", "Seafood_Length_Change","Seafood_Recommendation", 
                  "Priority_Score", "Breakeven Months"),
    caption = "Stores: Detailed Impact Analysis"
  )

```

```{r}
cat(paste("\nTotal stores analyzed:", nrow(final)))
cat(paste("\nStores with recommended changes:", sum(final$Priority_Score != "NO CHANGE")))
cat(paste("\nHigh priority changes:", sum(final$Priority_Score == "HIGH")))
```
